---
title: "Projects Guide"
author: "David Garcia <br><br> *TU Graz*"
date: "Foundations of Computational Social Systems"
output:
  xaringan::moon_reader:
    lib_dir: libs 
    css: [xaringan-themer.css, "libs/footer.css"]
    nature:
      beforeInit: ["libs/perc.js"]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---


```{r xaringan-themer, include=FALSE, warning=FALSE}
#This block contains the theme configuration for the CSS lab slides style
library(xaringanthemer)
library(showtext)
style_mono_accent(
  base_color = "#5c5c5c",
  text_font_size = "1.5rem",
  header_font_google = google_font("Arial"),
  text_font_google   = google_font("Arial", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

layout: true

<div class="my-footer"><span>David Garcia - Foundations of Computational Social Systems</span></div> 

---
#Projects for evaluation

The course assessment is based on a final course project. Exercise points are optional,the final project is **mandatory**.

Projects are done in groups that can have 
- at most 4 students 
- must mix students with different backgrounds (at least two different backgrounds per group).

The purpose of the project is to show what you have learned in the course in terms of data management, analysis, statistics, and interpretation.

Projects have to start from a research question and have an empirical focus, but at the same time critically reflect on methods, conclusions, and limitations.

---

#Projects timeline

The deadlines are up to the end of the day in Austria marked by each date.

- **19.12.2022:** Registration of projects.
- **03.02.2023, 14:00:** Project presentations session day 1.
- **17.02.2023:** Deadline for submission of final report. 

We offer a second date for project presentations and the submission of the final report for project teams that cannot make it on the first date:

- **03.03.2023, 14:00** Second date for project presentations session day 2.
- **17.03.2023** Second date for submission of final report.

You can get guidance for your project on two special online sessions with me:  
at 14:15 on 16.12.2022 and 27.01.2023.

---

# Project steps

**1. Form a group and choose a topic**
  - Start now! You can find partners on discord and around lectures. It can be coupled with Intro to CSS but it is not required.

**2. Project registration**. Deadline: 19.12.2022

**3. Project presentations**. On 03.02.2023 or 03.03.2023

**4. Submit the final report**. Deadline 17.02.2023 or 17.03.2023

---

# Project registration

- Register your project as a **plain text** email to dgarcia@tugraz.at. 
- Do not send it to css@tugraz.at, that account will not be monitored any more

A project registration is a short text that includes the following:

1. Project title
2. Names and email addresses of group members with a small note of disciplinary background of each
3. Date chosen for the presentation. If you don't choose a date, you will go for the first.
4. Research question(s)
5. Planned data retrieval and analysis to address the questions

---

## Important notes about registrations

.center[![:scale 25%](Warning.svg)]

###**Invest time on designing carefully your project registration!**

Once registered, you cannot change the research questions or project topic. 

You can later adapt steps in the planned data retrieval and analysis, but your final report has to argue why and justify the changes.

---

# Project presentations

- **5-minute** presentations about the state of results of their projects
- Students get questions from lecturers regarding their understanding of the topic and feedback on how to improve the project for the final report.
- Each presentation has to contain four slides:
  1. Research question and motivation
  2. Data and methods
  3. Results
  4. Conclusion and critique
- All members of the group have to participate in their presentation and answer questions.
- You can get guidance for your project on the designated sessions (16.12 and 27.01) or on discord up until the first project presentations date (03.02).  

**After 03.02.2023, no more guidance will be provided via discord or email.**

---
# Final report structure

1. **Motivation:** What question(s) do you seek to answer and why?

2. **Data retrieval:** Explain the interfaces or resources you used to collect all data necessary for the project.

3. **Data processing:** Explain how you filtered data, normalized values, computed additional variables, etc.

4. **Analysis:** Perform statistical analyses and visualizations that assess the question(s).

5. **Conclusion:** Evaluate answers to the question and their reliability.

6. **Critique:** Identify limitations and alternative explanations for your results.


---

# Submitting the final report

Send a final report as a PDF document (max. 6 pages, min. font size 11pt) via email to dgarcia@tugraz.at.

References do not count towards this 6-page limit. 

Projects can contain a link to a Github repository including the code to produce results, datasets if they can be shared, and additional figures or tables that can be referenced from the project report.

Plots should be correctly shown (named axes, visible scales) and writing has to be understandable. 

A good synthesis is much better than unnecessary information!
---

## Project grading

- 50% for the presentation
- 50% for the final report

Extra points are given when projects are based on open science principles (e.g. data and codes are available in a Github repository) and when data sources or methods go beyond what is covered in the course. 

**Projects do not need to report "positive results",** what is important is that you show how your have addressed your research questions, document any issues or deviations, and critically reflect on methods and results.

Remember that on top of the project presentation and report you can individually achieve an additional 4x5% (up to 20%) by handing in the exercise solutions.

---

# Three project paths

### A) Replicating a previous paper
  - Replication has to be critical and based on your own analysis, not the code of others
  - For simple papers, additional analysis beyond replication is expected

### B) Building on your Introduction to Computational Social Systems project

### C) Proposing your own idea

None of these paths has a higher expected grade, they are all equally valid.

---

# Project topic examples

- **Bias and fairness in recommender systems**

- **Integrating Survey Data and Digital Trails Data** 

- **Algorithm aversion as expressed on social media data** 

- **Social media polarization**

- **Popularity assortativity of musicians in Spotify**

- **Testing Social Impact Theory in the case of COVID19**

- **Sentiment about Donald Trump and income at the regional level**



